---
title: "Input to deliverable WP7"
author: "Michiel van Dijk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: ../Common/word_styles_01.docx
bibliography: ../Common/WP7.bib
    
---

```{r setup, include=FALSE}
library(pacman)
p_load(rprojroot)
root <- find_root(is_rstudio_project)

p_load(knitr)
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  dpi = 300,
  echo=FALSE, warning=FALSE, message=FALSE,
  fig.path = file.path(root,"Fig_tab_map/out-"),
  dev = "CairoPNG",
  dev.args = list(CairoPNG = list(bg = "transparent"))
  )

p_load(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

source(file.path(root, "Code/Fig2.R"))
#source(file.path(root, "Scripts/Tab.R"))
#source(file.path(root, "Scripts/Map.R"))
```

# Methodology [Might also go to the introduction]
A review of previous global modelling studies showed that results vary widely due to differences in (1) differences in scenario storylines and assumptions on drivers, (2) differences in the design of the models that are used to process the drivers and generate scenario outcomes, and (3) differences in the way and the extent to which results are reported [@VanDijk2014]. To overcome some of these problems, several initiatives have emerged that aim to harmonise assumptions on drivers and systematically compare results across an 'ensemble' of models. Examples of such an approach are the Agricultural Model Intercomparison Project (AgMIP) [@Rosenzweig2013, @Nelson2014c] and the recent quantification of Shared Socio-economic Pathways (SSPs) by the integrated assessment model (AIM) community [@VanVuuren2017]. These studies mainly focussed on the future pathways of global energy use and production, land use, emissions and the impact of climate change, with only limited attention for food security. None of the studies provides a comprehensive assessment of global food security going beyond basic indicators such as changes on prices and food production [@VonLampe2014]. Both AgMIP and the AIM community have used the five SSPs storylines and related quantification of drivers) to build their analysis [@ONeill2017]. Although, SSPs are well established and are now widely used by the research community for scenario analysis, they provide only one potential set of an infinite amount of futures. In this study we use four alternative narratives and drivers that were specifically developed with stakeholders for the assessment of global food security [@Dijk2016]. A model analysis of these scenarios provides and interesting alternative to the SSPs, to which they can be compared.

[ADD section on models]


# Food and Nutrition Security in the four worlds
## Global food and nutrition security
Figure X provides an overview of six key global food and nutrition security indicators that were produced by both models in the year 2050. In order to compare the various dimension of food security, the indicators were normalised using the following formula:

$$I_{it}=\frac {x_{it}-min(x_i)}{max(x_i)-min(x_i)}\times100$$

where $I_{it}$ is the normalised indicator, $x_{it}$ the value of indicator $i$ at time $t$, and $min(x_i)$ and $max(x_i)$ are the minimum and the maximum value of $xi$ across all scenarios over the base year (2010) and the final year (2050). This normalisation approach takes the evolution of indicators over time into account [@OECD2008]. If all values are larger than 0 it implies that the minimum value occurred in 2010 and, hence, there was an increase across all scenarios in comparison to the base year. Values of 100 on the outside of the diagram present more or higher values of the indicator and values of 0 reflect lower values. The purpose of the diagram is to present the overall picture on how the scenarios compare to each other in terms of food security conditions by showing their relative ranking. To obtain precise information on the pattern of growth over time and the change in levels, Figure 2 is presented.

Overall, the ECO scenario is located on the outside of the diagram for most indicators. At the global level and in comparison to the base year, the scenario is characterised by an increase in affordable food, higher calorie availability, a reduction in protein from animal origin and the highest share of calories from fruit and vegetables and high income per capita growth. The latter is lower than FFANF but because the scenario assumes a more equal distribution of income, it has a positive impact on global food security. A striking result is the low ranking on the share of nutritious food, which is defined as the share of food not spent on cereals. This is caused by: __[I do not understand the results for Calorie share on CER and VFN. Model results differ strongly and it is unclear why this is low for ECO. How is it possible for the MAGNET results that the share of CER is increasing and the share of VFN. Is this because the eat less meat? If so that would be a good explanation. For GLOBIOM, why is the share of VFN decreasing?]__. The TLTL forms the other extreme of the scenarios, illustrated by its location close to the centre of the diagram. Apart from share of nutritious food and reduction in protein of animal origin, it ranks lowest for all food security measures. The FFANF and ONEPW scenarios are in between the ECO and TLTL scenarios. In the FFANF world, income per capita growth is the highest but because of the impact of climate change, unsustainable behaviour and a diet dominated by the consumption of meat, food prices started to rise again at the end of this scenario. __[again a sentence is needed that explains the results for CALO_CER, which is highest in this scenario]__. Income per capita growth is relatively low in the ONEPW scenario, mainly because the very unequal distribution at the global level. As a consequence of high crop yield caused by technological change, food becomes more affordable. However, apart from the 1 percent elite, the majority of the global population __[Something about the diets]__. 

Figure X depicts the historical trend of the food security indicators (apart from share of calories from fruit and vegetables for which no data is available) and the model projections from 2010 till 2050. The solid line shows the model average and the shaded areas the bandwidth of both models. Overall, global food security is set to improve in most scenarios. On average, calorie availability and GDP per capita increase in al four scenarios, accompanied by a shift to more-nutritious food in all four worlds measured by a shift away from cereals in the diet. Food prices decrease in three out of four scenarios, TLTL being the exception. The other two indicators sketch a mixed picture. The supply of protein of animal origin, associated with a shift towards a diet with more meat is increasing in all worlds except for ECO. Although the consumption of more meat is less sustainable than other diets, it is likely to have a positive impact on food security in regions characterised by undernourishment, while the opposite is true for regions where obesity is a big problem. This issue is further addressed below where we discuss the results for the four major developing regions. The outcomes for share of calories from fruit and vegetables are not clear. The two models present completely contrasting results. __[CAN WE SAY MORE ABOUT WHY?]__.

```{r Fig-1, fig.width=8, fig.height=8}
Fig_radar_wld
```

##### Note: Lines represent the average of MAGNET-IMAGE and GLOBIOM outcomes. Food security indicators are normalised using the approach described in the text. 

__Comparison of trends in global scenario food security outcomes__

```{r Fig-2, fig.width=12, fig.height=12}
Fig_wld
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.

## Food and nutrition security in developing regions
The global picture belies a more detailed and nuanced regional picture. In this section, the six food security indicators along each of the four scenarios are reviewed for four major developing regions: Sub-Saharan Africa (SSA), South Asia (SASIA), East Asia (EASIA) and Latin America (LAC). As with the global picture, most indicators show an improvement across scenarios in all regions compared to the initial picture in 2010. There remains however a large degree of variation between the four worlds with overall greater improvements in food and nutrition security in the ECO and FFANF scenarios and lower improvements in ONEPW and TLTL. 

The results also show different development patterns across regions, which can broadly be divided in two groups. The first group is SSA and SASIA, which make up the relatively least developed countries. These countries are characterised by relatively low calorie availability (around 2,500 kcal/cap/day and income per capita (around 30,00-5,000 2011 PPP USD/cap), and a diet that consists of limited meat consumption (16-25 gram/cap/day) in 2010. In these two regions, food security improves markedly in the ECO, ONEPW and FFANF scenarios. An increase in income per capital results in better access to food, while at the same time calorie availably increases to more than 3000 kcal/cap/day.  In SSA this is accompanied by a decrease in food prices, while in SASIA food prices rise slightly. Both regions also show a similar shift in the diet. In both SSA and SASIA, people start to eat more meat, illustrated by the increase supply of protein from animal origin. The increase in welfare in SSA leads to a very strong increase in meat consumption in the FFANF scenario, while this is much more modest in ECO, where sustainability plays a key role in consumers choices. Nonetheless, the supply of protein from animal origin in SSA and SASIA in 2050 remains still very low in absolute levels in comparison to the rich countries today (around 50 grams/cap/day). In both regions the share of energy supply from cereals decreases slightly __[CAN WE EXPLAIN WHY?]__ while the share of calories from fruit and vegetables increases (SSA) or also decreases somewhat (SASIA). However, the latter result should be treated with caution as the models tend to disagree about the projections for fruit and vegetables consumption. In contrast to the other three scenarios, food security is not or only very moderately improving in the TLTL world in SSA and SASIA. Income per capita, calorie availability and diets improve only slightly, while at the same time access to food deteriorates due to steeply rising food prices. 

In EASIA and LAC, a somewhat different pattern can be observed  In these two regions food security is already relatively much better in 2010 in comparison to SSA and SASIA. Calorie availability (around 3000 kcal/cap/day), income (12,000-13,000 2011 PPP USD/cap) and meat consumption (26-37 grams/cap/day) are substantially higher. Strikingly, differences in the share of dietary energy supply derived from cereals are much less pronounced between both groups of countries (around 60% for SSA and SASIA in comparison to 40-50% for EASIA and LAC). Also in EASIA and LAC, food security improves in the ECO, ONEPW and FFANF scenarios but improvements are less dramatic than in the other two regions because of the high base levels. Calorie availability increases to around 3,250 kcal per cap in both regions, which is __[can we argue that this is more than what is needed, reaching obesity levels?]__. Equally, the consumption of meat increases to around 45-50 grams/cap/day, which is comparable to rich countries. As basic needs are already fulfilled and relative income is high, consumers make the decision to have more sustainable diet towards less meat in the ECO scenario. Food prices slightly increase in EASIA and decrease in LAC. This is, however, expected to have limited impact on food security as consumers spend a relatively low share of their (high) income per capital on food items. __[Maybe a sentence can be added on the cereal and fruit diet part but I find them difficult to explain]__



__SSA__

```{r Fig-3}
Fig_SSA
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.

__EASIA__

```{r Fig-4}
Fig_EASIA
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.


__SASIA__

```{r Fig-5}
Fig_SASIA
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.

__LAC__

```{r Fig-6}
Fig_LAC
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.




# Discussion [Perhaps not for now but for scientific article]
Here we should probably say something about:
- Why models differ, going (a bit) into difference PE and CGE models, as well as elasticities etc
- Compare with SSP outcomes if possible.





# References