---
title: "Input to deliverable WP7"
author: "Michiel van Dijk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: ../Common/word_styles_01.docx
bibliography: ../Common/WP7.bib
    
---

```{r setup, include=FALSE}
library(pacman)
p_load(rprojroot)
root <- find_root(is_rstudio_project)

p_load(knitr)
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  dpi = 300,
  echo=FALSE, warning=FALSE, message=FALSE,
  fig.path = file.path(root,"Fig_tab_map/out-"),
  dev = "CairoPNG",
  dev.args = list(CairoPNG = list(bg = "transparent"))
  )

p_load(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

source(file.path(root, "Code/Fig2.R"))
#source(file.path(root, "Scripts/Tab.R"))
#source(file.path(root, "Scripts/Map.R"))
```

# Methodology [Might also go to the introduction]
A review of previous global modelling studies showed that results vary widely due to differences in (1) differences in scenario storylines and assumptions on drivers, (2) differences in the design of the models that are used to process the drivers and generate scenario outcomes, and (3) differences in the way and the extent to which results are reported [@VanDijk2014]. To overcome some of these problems, several initiatives have emerged that aim to harmonise assumptions on drivers and systematically compare results across an 'ensemble' of models. Examples of such an approach are the Agricultural Model Intercomparison Project (AgMIP) [@Rosenzweig2013, @Nelson2014c] and the recent quantification of Shared Socio-economic Pathways (SSPs) by the integrated assessment model (AIM) community [@VanVuuren2017]. These studies mainly focussed on the future pathways of global energy use and production, land use, emissions and the impact of climate change, with only limited attention for food security. None of the studies provides a comprehensive assessment of global food security going beyond basic indicators such as changes on prices and food production [@VonLampe2014]. Both AgMIP and the AIM community have used the five SSPs storylines and related quantification of drivers) to build their analysis [@ONeill2017]. Although, SSPs are well established and are now widely used by the research community for scenario analysis, they provide only one potential set of an infinite amount of futures. In this study we use four alternative narratives and drivers that were specifically developed with stakeholders for the assessment of global food security [@Dijk2016]. A model analysis of these scenarios provides and interesting alternative to the SSPs, to which they can be compared.

[ADD section on models]


# Food and Nutrition Security in the four worlds
## Global food and nutrition security
Figure X provides an overview of six key global food and nutrition security indicators that were produced by both models in the year 2050. In order to compare the various dimension of food security, the indicators were normalised using the following formula:

$$I_{it}=\frac {x_{it}-min(x_i)}{max(x_i)-min(x_i)}\times100$$

where $I_{it}$ is the normalised indicator, $x_{it}$ the value of indicator $i$ at time $t$, and $min(x_i)$ and $max(x_i)$ are the minimum and the maximum value of $xi$ across all scenarios over the base year (2010) and the final year (2050). This normalisation approach takes the evolution of indicators over time into account [@OECD2008]. If all values are larger than 0 it implies that the minimum value occured in 2010 and, hence, there was an increase across all scenarios in comparison to the base year. Values of 100 on the outside of the diagram present more or higher values of the indicator and values of 0 reflect lower values. The purpuse of the diagram is to present the overall picture on how the scenarios compare to each other in terms of food security conditions by showing their relative ranking. To obtain precise information on the pattern of growth over time and the change in levels, Figure 2 is presented.

Overall, the ECO scenario is located on the outside of the diagram for most indicators. At the global level and in comparison to the base year, the scenario is characterised by an increase in affordible food, higher calorie availability, a reduction in protein from animal origin and the highest share of calories from fruit and vegetables and high income per capita growth. The latter is lower than FFANF but because the scenario assumes a more equal distribution of income, it has a positive impact on global food security. A striking result is the low ranking on the share of nutritious food, which is defined as the share of food not spent on cereals. This is caused by: __[I do not understand the results for Calorie share on CER and VFN. Model results differ strongly and it is unclear why this is low for ECO. How is it possible for the MAGNET results that the share of CER is increasing and the share of VFN. Is this because the eat less meat? If so that would be a good explanation. For GLOBIOM, why is the share of VFN decreasing?]__. The TLTL forms the other extreme of the scenarios, illustrated by its location close to the centre of the diagram. Apart from share of nutritious food and reduction in protein of animal origin, it ranks lowest for all food security measures. The FFANF and ONEPW scenarios are in betweeen the ECO and TLTL scenarios. In the FFANF world, income per capita growth is the highest but because of the impact of climate change, unsustainable behaviour and a diet dominated by the consumption of meat, food prices started to rise again at the end of this scenario. __[again a sentence is needed that explains the results for CALO_CER, which is highest in this scenario]__. Income per capita growth is relatively low in the ONEPW scenario, mainly because the very unequal distribution at the global level. As a consequence of high crop yield caused by technological change, food becomes more affordable. However, apart from the 1 percent elite, the majority of the global population __[Something about the diets]__. 

Figure X depicts the historical trend of the food security indicators (apart from share of calories from fruit and vegetables for which no data is available) and the model projections from 2010 till 2050. The solid line shows the model average and the shaded areas the bandwidth of both models. Overall, global food security is set to improve in most scenarios. On average, calorie availability and GDP per capita increase in al four scenarios, accompanied by a shift to more-nutrious food in all four worlds measured by a shift away from cereals in the diet. Food prices decrease in three out of four scenarios, TLTL being the exception. The other two indicators sketch a mixed picture. The supply of protein of animal origin, associated with a shift towards a diet with more meat is increasing in all worlds except for ECO. Although the consumption of more meat is less sustainable than other diets, it is likely to have a positive impact on food security in regions characterised by undernourishment, while the opposite is true for regions where obisity is a big problem. This issue is further addressed below where we discuss the results for the four major developing regions. The outcomes for share of calories from fruit and vegetabels are not clear. The two models present completely contrasting results. __[CAN WE SAY MORE ABOUT WHY?]__.






```{r Fig-1, fig.width=8, fig.height=8}
Fig_radar_wld
```

##### Note: Lines represent the average of MAGNET-IMAGE and GLOBIOM outcomes. Food security indicators are normalised using the approach described in the text. 

__Comparison of trends in global scenario food security outcomes__

```{r Fig-2, fig.width=12, fig.height=12}
Fig_wld
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.

## Food and nutrition security in developing regions
The global picture belies a more detailed and nuanced regional picture. The development of  food and nutrition security in Africa for each of the four worlds is shown in Table 6. As with the global picture, only a few indicators show a worsening compared to the initial picture in 2010. There remains however a large degree of variation between the four worlds with greater improvements in food and nutrition security in the ECO and FFANF scenarios and lower improvements in ONEPW and TLTL.

The quantity of food available, measured in calories per capita per day, is highest in the ECO scenario in Africa. There are some improvements in diet quality as the amount meat available and the share of fruit and vegetables in the average diet improve, although this is slightly offset by an increase in the share of calories from cereals. The share of food expenditure in total expenditure improves dramatically, suggesting large improvements in food access; driven by higher incomes and lower prices. Africa is much less reliant on cereal imports in this world by 2050 and significantly more able to finance food imports from export revenue. These indicators point to an increase in the production of cereals within Africa, partly encouraged by higher levels of trade restrictions in this scenario. What’s happening with production? More exports?

The quantity of calories available for the average consumer is also good in the FFANF scenario. The diet quality indicators suggest improving diets with fewer calories from cereals and more from meat and vegetables; although slightly lower in vegetables that in ECO and ONEPW. The largest fall in the share of food expenditure in total expenditure is seen in this scenario, suggesting the largest improvement in food access due to high growth and medium-high reductions in food prices. Africa is increasingly able to afford food imports in this future world; to a similar extent as in the ECO scenario. However, the trade liberalisation in this scenario improves access to imports from other regions and keeps Africa relatively reliant on cereal imports. # Initial position Talk about structural change here – producing non cereals that find their markets abroad?

All indicators of FNS improve in the ONEPW scenario but the improvement is less than in the equal scenarios of ECO and FFANF. The increase in calories available is slightly lower than the leading scenarios., however diet quality improves as the share of calories from cereals falls, the share from vegetables increases and the availability of protein increases. Not consistent with non-elite narrative. Can all shares increase? What’s decreasing? The share of food expenditure falls by much less than in the equal scenarios, despite falling food prices, due to slower growth in per capita incomes. Lower growth also contributes to a smaller improvement in the ability of the region to pay for its food imports from export revenue. Despite trade liberalisation, Africa becomes more self-sufficient in cereals in this world.

GDP growth in Sub-Saharan Africa falls into two groups: high growth in FFANF and ECO and lower growth in ONEPW and TLTL. Higher population growth in the low growth scenarios exacerbates the difference leading to high per capita GDP in the equal scenarios and low per capita GDP in the unequal scenarios. This is consistent with the narrative of the equal worlds where there is convergence between low-medium income regions and high income regions. 
Food and nutrition security improves the least in the Too Little, Too Late world. This scenario shows the lowest levels of calories available of the four future worlds. The share of calories from cereals increase and actually decrease from fruit and vegetables suggesting worsening diet quality, despite an increase in protein availability. TLTL is the only one of the four worlds to project food price increases by 2050, which together with low income growth, leads to the lowest improvement in the share of food expenditure in total household expenditure. The difficult economic circumstances that are characteristic of this scenario, play out in an increase in Africa’s dependence on cereals imports in a world of increasingly restrictive, and costly, trade. The ability of African to finance food imports from merchandise exports improves by less than in the equal scenarios, although the least improvement is still seen in the ONEPW world.

Food prices rise in the TLTL world due relatively high population growth, coupled with lower levels of  technical change and yield growth. Population growth is similar in ONEPW and TLTL, however the efficiency gains present in the ONEPW world, mean that resources are growing (through technical change), faster than the demand from the additional population is rising and prices fall, despite the additional pressure on land arising from the land protection policy. The limited technological change in TLTL means that the demand pressure builds up and causes prices to rise. The introduction of reduced meat consumption has little impact in Africa as the reductions are assumed to take place in high income countries.



Per capita GDP growth also falls into two groups in South Asia with high growth per capita in the equal scenarios and lower growth per capita in the unequal scenarios. Higher population growth in the unequal scenarios exacerbate the differences by sharing the available economic growth among more people.
The equal scenarios have the highest per capita growth rates in East Asia and the unequal growth rates have the lowest, but the difference between the equal and unequal scenarios is less pronounced in this region. Almost all of the variation in per capita GDP growth across the four worlds comes from differences in the headline growth rates as population growth is similar in all worlds, with just a slightly higher rate in ECO which brings down the GDP per capita rate.





__SSA__

```{r Fig-3}
Fig_SSA
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.

__EASIA__

```{r Fig-4}
Fig_EASIA
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.


__SASIA__

```{r Fig-5}
Fig_SASIA
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.

__LAC__

```{r Fig-6}
Fig_LAC
```

##### Source: Historical price information from World Bank. Historical calorie availability from FAOSTAT. All other historical data from FAO food security indicators.




# Discussion [Perhaps not for now but for scientific article]
Here we should probably say something about:
- Why models differ, going (a bit) into difference PE and CGE models, as well as elasticities etc
- Compare with SSP outcomes if possible.





# References